---
title: "simpleRegress"
author: "Caitlyn Hense"
date: "4/17/25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model WEIGHT ~ NECK  this is an example  

```{r}
simp <- lm(WEIGHT~NECK,data=bears)
plot(WEIGHT~NECK,data=bears)
abline(simp)
summary.lm(simp)
```



## Compare all models that predict WEIGHT other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 
```{r}
mHEADLEN <- lm(WEIGHT~HEADLEN,data=bears)
plot(WEIGHT~HEADLEN,data=bears)
abline(mHEADLEN)
summary.lm(mHEADLEN)
```

From our observations, the residual standard error is 67.81 and our adjusted R-squared is 0.69. Ordering the models from best to worst, the best one is the weight explained by neck followed by the weight explained by HEADLEN.

```{r}
mHEADWTH <- lm(WEIGHT~HEADWTH,data=bears)
plot(WEIGHT~HEADWTH,data=bears)
abline(mHEADLEN)
summary.lm(mHEADWTH)
```

Best to Worst ordered by RSE  
Neck RSE is 43.91   
HEADLEN RSE is 67.81  
HEDWTH RSE is 76.41  

```{r}
LENGTH <- lm(WEIGHT~LENGTH,data=bears)
plot(WEIGHT~LENGTH,data=bears)
abline(LENGTH)
summary.lm(LENGTH)
```

Best to Worst ordered by RSE  
Neck RSE is 43.91  
LENGTH RSE is 61.83  
HEADLEN RSE is 67.81  
HEDWTH RSE is 76.41 
 
```{r}
CHEST <- lm(WEIGHT~CHEST,data=bears)
plot(WEIGHT~CHEST,data=bears)
abline(CHEST)
summary.lm(CHEST)
```

Best to Worst ordered by RSE  
CHEST RSE is 33.08  
Neck RSE is 43.91  
LENGTH RSE is 61.83   
HEADLEN RSE is 67.81  
HEDWTH RSE is 76.41 
 



